---
title: "PSTAT131/HW2"
author: "Cynthia Cao"
date: "2022-10-08"
output: pdf_document
---

```{r}
library(tidyverse)
library(tidymodels)
abalone <- read_csv("/Users/cynnnthiaaa/Desktop/abalone.csv")
```

Q1:
```{r}
abalone$age <- abalone$rings+1.5
abalone
ggplot(abalone, aes(x=age)) + geom_histogram()
```
From the graph, we can see that the distribution of age is roughly normal, most observations are between age of 7- 15.



Q2:
```{r}
set.seed(3456)
abalone_split <- initial_split(abalone, prop = 0.75,strata = age)
train  <- training(abalone_split)
test   <- testing(abalone_split)
train
test
```

Q3:
```{r}
new_abalone <- select(train,-rings)
abalone_recipe <- recipe(age~.,data = new_abalone) %>%
  step_dummy(type)

abalone_recipe <- step_interact(abalone_recipe, terms = ~ shucked_weight : starts_with('type'))
abalone_recipe <- step_interact(abalone_recipe, terms = ~ longest_shell : diameter)
abalone_recipe <- step_interact(abalone_recipe, terms = ~ shucked_weight : shell_weight)

abalone_recipe <- step_center(abalone_recipe, longest_shell, diameter, height, whole_weight, shucked_weight, viscera_weight, shell_weight)
abalone_recipe <- step_scale(abalone_recipe, longest_shell, diameter, height, whole_weight, shucked_weight, viscera_weight, shell_weight)

abalone_recipe
```
Since we already know the condition that age=rings+1.5, the relationship between two variables is fixed, then its unnecessary to include it in the prediction. 


Q4:
```{r}
lm_model <- linear_reg() %>% 
            set_engine('lm') %>% 
            set_mode('regression')
```

Q5:
```{r}
workflow1 <- workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(abalone_recipe)
```

Q6:
```{r}
fit1 <- fit(workflow1, new_abalone)
type <- c('F')
longest_shell <- c(0.50)
diameter <- c(0.10)
height <- c(0.30)
whole_weight <- c(4)
shucked_weight <- c(1)
viscera_weight <- c(2)
shell_weight <- c(1)
data1 <- data.frame(type, longest_shell,diameter,height,whole_weight,
                         shucked_weight, viscera_weight, shell_weight)
predict(fit1, new_data = data1)
```
Q7:
```{r}
library(yardstick)
metric <- metric_set(rsq, rmse, mae)
result1 <- predict(fit1, new_data=new_abalone %>% select(-age))
result2 <- bind_cols(result1, new_abalone %>% select(age))
result2
metric(result2, truth = age, estimate = .pred)
```
R Square: 0.5580969	 (about 55.81% of variability of Y can be explained by X)
Root Mean Square Error: 2.1463035	
Mean Absolute Error: 1.5375481	


